---
title: "arm_b_5p1m_KD.qmd"
author: "Krish Desai"
format: html
editor: visual
---

## **Reference Download and Preparation**

```{R}
#Address
URL <-"https://ftp.ncbi.nlm.nih.gov/genomes/refseq/fungi/Saccharomyces_cerevisiae/latest_assembly_versions/GCF_000146045.2_R64/GCF_000146045.2_R64_genomic.fna.gz"

#Destenatoin file (output)

dest_file <- "GCF_000146045.2_R64_genomic.test.fna.gz"

#Download the file with wget
download.file(URL, method = "wget", destfile = dest_file)

con <- gzfile("GCF_000146045.2_R64_genomic.test.fna.gz", "rt")
ygen <- readLines(con)
close(con)

# 2. Define the new UCSC-style headers (chrI through chrXVI + chrM)
new_headers <- c(paste0(">chr", as.roman(1:16)), ">chrM")

# 3. Find and Replace
header_lines <- grep(">", ygen)
#Look at the header for chromosome 1
header_lines[1]

# Safety Check: Yeast has 16 nuclear chromosomes + 1 mitochondrial
if(length(header_lines) == 17) {
  ygen[header_lines] <- new_headers
  writeLines(ygen, "S288C_UCSC.fna")
  message("SUCCESS: Reference renamed to UCSC standards.")
} else {
  stop("Error: Chromosome count mismatch! Check your input file.")
}
```

# **Part 2: The Reconstruction (Bash)**

## **Step 1: Assembly with Flye & Log Audit**

```{bash}
#| echo: False
#| eval: False


# Activate conda evnvironment Fly_env
Conda activate flye_env

#Crete genome assembly folder
mkdir data/Genome_assembly

#Run fly on your daa
flye /Users/kdesai01/Desktop/Genmics_and_transcriptomics/Week2_Yeast_Project/data/clean_reads/ERR1539001_Clean.fastq.gz --out-dir data/assamble_genomes --genome-size 12m --threads 8  --scaffold --iterations 5
```

## **Step 2: Polishing with Medaka**

```{bash}
#| echo: False
#| eval: False


# Deactivate conda evnvironment Fly_env
Conda deactivate

# Activate conda evnvironment medaka_env
Conda activate medaka_env

#Crete genome assembly folder
mkdir data/apolished_assembly

#Run fly on your daa
medaka_consensus  -i /Users/kdesai01/Desktop/Genmics_and_transcriptomics/Week2_Yeast_Project/data/clean_reads/ERR1539001_Clean.fastq.gz -d /data/assamble_genomes/assembly.fasta -o data/apolished_assembly -t 8 
```

# **Part 3: Technical & Biological Audit (Docker)**

## **1. QUAST (Technical Quality)**

```{bash}
#| echo: False
#| eval: False

echo 'export PATH="$HOME/.docker/bin:$PATH"' >> ~/.zshrc
source ~/.zshrc

docker pull quay.io/biocontainers/quast:5.3.0--py313pl5321h5ca1c30_2

docker run --rm quay.io/biocontainers/quast:5.3.0--py313pl5321h5ca1c30_2 quast.py -h

docker run --rm -v "$(pwd)":/data -w /data quay.io/biocontainers/quast:5.3.0--py313pl5321h5ca1c30_2  quast.py data/apolished_assembly/consensus.fasta --fungus -r S288C_UCSC.fna -o reports/quast_output
```

## **2. BUSCO (Biological Completeness)**

```{bash}
docker pull busco/busco:v6.0.0_cv1

docker run --rm \
-v "$PWD":/data \
-w /data \
ezlabgva/busco:v6.0.0_cv1 \
busco -i data/apolished_assembly/consensus.fasta  -o reports/busco_repots -l saccharomycetaceae_odb12 -m genome -c 8
```

# **Part 4: Structural Validation (R & Gviz)**

## **Step 1: High-Efficiency Alignment & Indexing**

```{bash}
#ctivate mapping environment
conda deactivate

conda create -n mapping_env -c bioconda minimap2 samtools -y
conda activate mapping_env
minimap2 -h
minimap2 -h
# Map and Sort in one step
# minimap2 -a (output SAM) | samtools sort (input from pipe)
minimap2 -a -x asm5 S288C_UCSC.fna data/apolished_assembly/consensus.fasta | samtools sort -o assembly_mapped.sorted.bam -

# Create the index
samtools index assembly_mapped.sorted.bam

# Run the coverage report
samtools coverage assembly_mapped.sorted.bam -o assembly_cov.txt
```

```{r}
# Read the coverage report
assembly_cov <- read.table("assembly_cov_KD.txt", sep = "\t", header = TRUE, comment.char = "")
print(assembly_cov)

```

## **Step 2: Programmatic Discovery & Selection**

```{r}
#| echo: False
#| eval: False

#Loading the Engine
library(Gviz)
library(Rsamtools)
library(GenomicAlignments)
library(GenomicRanges)

# 1. Audit the Stats: Find chromosomes > 500kb with high density
stats <- idxstatsBam("assembly_mapped.sorted_KD.bam")
print(stats)

# 2. Extract ranges for the entire genome
bam_file <- "assembly_mapped.sorted_KD.bam"
alignment_ranges <- bam_file |> readGAlignments() |> GRanges()

# Look at the ranges for ALL chromosomes to see your global success
range(alignment_ranges)

# Peek at the individual read boundaries
ranges(alignment_ranges)
```

## **Step 3: The Custom Visual Audit**

```{r}
library(Gviz)

# 1. Create the Data Track using the GRanges object (The "Scissors" approach)
align_track <- AlignmentsTrack(alignment_ranges, name = "Assembly Alignments", fill = "darkblue")

# 2. Create the Ruler
axis_track <- GenomeAxisTrack()

# 3. Plot
# Replace the chromosome name and numbers with the ones you discovered!
plotTracks(list(axis_track, align_track), 
          from = 81007, 
           to = 195324, 
           chromosome = "chr1")
```
